---
title: "M234-lab6"
author: "Zixi Zhang"
date: "2023-02-24"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Users/bruce/Documents/Biostat.courses/23W-M234/234Lab/Lab6")
library(R2jags)
load("AddBurnin.RData")
load("lab5data.RData")
library(lattice)
library(ggplot2)
```


## 1.	Print model 1 and on the printout, identify the following:

```{r}
# sink("model1.txt")
# cat("
# # Response is number of Exencephaly malformations 
# # out of the total number of live fetuses
# # in the litter.  
# # Model #1.  
# model
# {
#         for (i in 1:N)
#         { 
#                 y[i] ~ dbin(p[i], n[i])
#                 logit(p[i]) <-  alpha[1] +alpha[2]*(z[i]-695) + beta[i] + 
#                                 x1[i]*alpha[3] +x2[i]*alpha[4]
#                 beta[i] ~ dnorm(0, tau)
#         }
        # for (j in 1:4)
        # {
        #         alpha[j] ~dnorm(0, 0.001)
        # }
        # sigsq ~ dunif(0,6)
        # tau <- 1/sigsq
        # sigma <- sqrt(sigsq)
# 	for(j in 1:npred)
# 		{
# 			lppred[j] <- alpha[1] + alpha[2]*(pred[j]-695) + betanew
# 			logit(probpred[j]) <- lppred[j]
# 		}
# 		betanew ~ dnorm(0,tau)
# }
#     ",fill = TRUE)
# sink()

```

### a.	Where the likelihood is modeled;

logit(p[i]) <-  alpha[1] +alpha[2]*(z[i]-695) + beta[i] +  x1[i]*alpha[3] +x2[i]*alpha[4]

### b.	Where the priors are modeled;

for (j in 1:4)
{
    alpha[j] ~dnorm(0, 0.001)
}
sigsq ~ dunif(0,6)
tau <- 1/sigsq
sigma <- sqrt(sigsq)

### c.	Where the linear predictor is specified;

lppred[j] <- alpha[1] + alpha[2]*(pred[j]-695) + betanew

### d.	Where the predictions are modeled;

	for(j in 1:npred)
		{
			lppred[j] <- alpha[1] + alpha[2]*(pred[j]-695) + betanew
			logit(probpred[j]) <- lppred[j]
		}
		betanew ~ dnorm(0,tau)
}

### e.	Where what predictions we are interested in are given.  

logit(probpred[j]) <- lppred[j]

## 2.	Give time series plots and autocorrelation plots for sigma and for alpha[2].  Give an estimate of how many iterations are sufficient for good estimation.  
```{r}
model1parameters= c("alpha", "p", "probpred", "beta", "tau", "sigma")

#### Did you load the data yet???

run1=jags(model1data, model1inits, model1parameters, "model1.txt", n.chains=3, n.iter=11000, n.burnin=0, n.thin=1, DIC=F)

Output1=AddBurnin(run1$BUGSoutput$sims.array,burnin=1000,n.thin=1)

# print the summary and see order of the parameters 
# print(Output1$Burnin.Summary)
```

```{r}
par(mfrow=c(1,2))
acf(Output1$Burnin.sims.array[,1,2], main="", lag.max=150) 
mtext("alpha[2]",side=3, line=1, cex=.8)
acf(Output1$Burnin.sims.array[,1,137], main="",lag.max=150) 
mtext("sigma", side=3, cex=.8, line=1)
```

100 iterations are sufficient for good estimation.


## 3.	Which p[i] is largest?  Discuss (1-2 sentences) how you decide this, as there is more than one way to determine this.

```{r}
#Boxplots for p
par(mfrow=c(1,1)) # sets plots in a 1x1 grid
#You may need to change the window size of the plot
boxplot(Output1$Burnin.sims.matrix[,c(seq(67,128))], range=0, axes=F)
axis(side=1,at=seq(1,62,5), labels=paste("p",seq(1,62,5)))
axis(side=2, at=0:5/5, labels=0:5/5)
box()
```

From the plot, we can see p[25] is the largest.
Here is another way.
```{r}
names <- paste0("p[", 1:62, "]")
p <- as.data.frame(Output1$Burnin.Summary[names, ])
max_row <- p[which.max(p$mu.vect),]
max_index <- rownames(max_row)
max_index
```


## 4.	Which beta[i] is largest?  Does it correspond to the same case as in Question 2?  Explain: Do the largest beta[i] and the largest p[i] have to correspond to the same case [i]?  

```{r}
#Boxplots for Beta
boxplot(Output1$Burnin.sims.matrix[,c(seq(5,66))], range=0, axes=F)
axis(side=1,at=seq(1,62,5), labels=paste("Beta",seq(1,62,5)))
axis(side=2, at=c(-5,0,5), labels=c(-5,0,5))
box()
```

From the plot, we can see Beta[29] is the largest. It does not correspond to the same case as in Question 2?  The largest beta[i] and the largest p[i] doe not have to correspond to the same case [i].

## 5.	Plot posterior estimates of (.05, median, .95) for each prediction against blood glucose level.  Connect the dots between neighboring glucose levels.  Label and document your plot carefully.  

```{r}
plotdf1 <-as.matrix(run1$BUGSoutput$summary)
probnames <- c("probpred[1]","probpred[2]","probpred[3]","probpred[4]","probpred[5]","probpred[6]","probpred[7]","probpred[8]")
plotdf2 <- plotdf1[probnames, c("2.5%", "50%", "97.5%")]
plotdf <- data.frame(bloodglucose = seq(100, 800, 100), 
                 label = rep(c("2.5%", "Median", "97.5%"), each = 8), 
                 prediction = c(plotdf2[, 1], plotdf2[, 2], plotdf2[, 3]))
```

```{r}
ggplot(data = plotdf, aes(x = bloodglucose, y = prediction, colour = label)) +
  geom_point() +
  geom_line() +
  labs(title = "Plot of posterior estimates of  prediction",
       x = "blood glucose level", y = "prediction prob.") +  
  ylim(0, 1) + theme_bw()
```
## 6.	Repeat questions 1-4 for model.  One question does not make sense.  Which is it?  Why doesn’t it make sense?  
```{r}
# sink("model2.txt")
# cat("
# # Model #2  
# model
# {
#         for (i in 1:N)
#         { 
#                 y[i] ~ dbin(p[i], n[i])
#                 p[i] ~ dbeta(a[i],b)
#                 a[i] <- b * exp(alpha[1] +alpha[2]*(z[i]-695) + 
#                         x1[i]*alpha[3] +x2[i]*alpha[4])
#         }
        # for (j in 1:4)
        # {
        #         alpha[j] ~dnorm(0, 0.001)
        # }
        # b ~ dgamma(ba,bb)
# }
#     ",fill = TRUE)
# sink()
```

### a.	Where the likelihood is modeled;

p[i] ~ dbeta(a[i],b)

### b.	Where the priors are modeled;

for (j in 1:4)
{
    alpha[j] ~dnorm(0, 0.001)
}
b ~ dgamma(ba,bb)

### c.	Where the linear predictor is specified;

a[i] <- b * exp(alpha[1] +alpha[2]*(z[i]-695) +  x1[i]*alpha[3] +x2[i]*alpha[4])

### d.	Where the predictions are modeled;
### e.	Where what predictions we are interested in are given.  

p[i]

Questions 2 and 4 do not make sense. Because model 2 does not have sigma and beta.

```{r}
# run1=jags(model2data, model2inits, model1parameters, "model1.txt", n.chains=3, n.iter=11000, n.burnin=0, n.thin=1, DIC=F)
# 
# Output1=AddBurnin(run1$BUGSoutput$sims.array,burnin=1000,n.thin=1)
```




